State of the Practice for Lattice Boltzmann Method Software

### 2022-04-26 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Revisions to Section 10.

\subsection{CI Usage Compared to Research Software in General}

@@

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are suggestions for consideration in the future. We
will not be repeating previously discussed ideas here, such as CI, documentation
of APIs, etc.  Our aim is to mention ideas that are at least somewhat beyond
conventional best practices. The ideas listed here have the potential to become
best practices in the medium to long-term. The ideas below are roughly in the
order of increasing novelty.

@@

Except for \citet{ThielEtAl2020}, the research software guidelines that we
consulted do not mention linters.  However, we believe that linters have the potential
to improve code quality at a relatively low cost.  A linter is a tool that
statically analyzes code to find programming errors, stylistic inconsistencies
and suspicious constructs \citep{Wikipedia2022_Lint}. Linters can be used to
spot check code files, or even better as part of a continuous integration
system, as discussed in Section~\ref{Sec_CompareTools}.

@@

Linters can address the pain points for LBM developers.  For instance, a linter
can decrease the amount of development time (\ppref{P_LackDevTime}) by
decreasing the number of mundane mistakes programmers have to catch.  Since the
linter can include rules that capture the wisdom of senior programmers, it can
help newer developers avoid common mistakes (\ppref{P_LackSoftDevExp}).
Although there do not appear to be many linters for checking parallel
programming mistakes (\ppref{P_TechnologyHurdles}), Parallel Lint is an option
for OpenMP. Consistently using a linter can guard against some technical debt
(\ppref{P_TechDebt}). Although linters are tools for code analysis, similar
ideas can be applied to ``lint'' documentation to ensure adherence to basic
rules.  The use of tools to check documentation is a partial explanation for the
relatively higher quality of statistical tools that are part of the
Comprehensive R Archive Network (CRAN) \citep{SmithEtAl2018_StatSoft}.

@@

Peer review is a strategy already adopted by some LBM developers to improve
confidence in their software (Section~\ref{SecReproducibility}), but the
benefits of peer review can be improved by making its application more rigorous.
\citet{Jones2008} shows that rigorous inspection finds 60-65\% of latent defects
on average, and often tops 85\% in defect removal efficiency.  By way of
comparison, most forms of testing average between 30 and 35\% for defect removal
efficiency \citep{EbertAndJones2009, Jones2008}.  A formal code inspection
involves asking the reviewer to either follow a review checklist (check
consistency of variable names, look for terminating loops, etc.), or perform
specific review tasks (like summarize the purpose of the code, create a data
dictionary for the module assigned, cross-reference the code to the technical
manual, etc.) The task based inspection approach has been effectively used for
research software, as described by \citet{KellyAndShepard2000}. Task based
inspection is an ideal fit with an issue tracking system, like GitHub.  The
review tasks can be issues, so that they can be easily assigned, monitored and
recorded.  Other potential issues for the tracker include assigning junior team
members to test installation instructions and getting-started tutorials.

Rigorous peer review addresses the same pain points as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_LackSoftDevExp},
\ppref{P_TechnologyHurdles}, \ppref{P_TechDebt} and \ppref{P_Correctness}. Peer
review addresses these pain points by efficiently searching for defects and
problems.

@@

To address the pain point of a lack of funding (\ppref{P_LackFunding}), LBM
developers should watch for new opportunities to publish their research software
source code. For instance, the Journal of Open Source Software (JOSS) reviews
and publishes articles that credit the scholarship contained in the software
itself \citep{SmithEtAl2018-joss}.  \citet{SmithEtAl2016-softcite} presents a
set of software citation principles that may encourage broad adoption of a
consistent policy for software citation across disciplines and venues.
\citet{ChueHongEtAl2019} provides a software citation checklist for developers
to make a release of their software citable.  \citet{KatzEtAl2021} provides
further guidance on software citation.  Another option is to use GitHub for
citation, since it is relatively easy to add metadata to repositories and to
generate citations for those repositories \citet{Smith2021-CitationsOnGitHub}.

@@

To address the pain point of ensuring correctness (\ppref{P_Correctness}),
Section~\ref{SecReproducibility} outlined techniques that LBM developers use to
circumvent the oracle problem for testing.  Additional techniques to build
system tests are as follows \citep{Smith2016}:

@@

As discussed in Section~\ref{Sec_CompareArtifacts}, developers rarely write requirements specifications for research software. Although a full requirements
specification has advantages \citep{SmithEtAl2007, SmithAndLai2005},
given the lack of development time (\ppref{P_LackDevTime}), we recommend the
intermediate step of augmenting the existing theory manuals with some
requirements information.  Table~\ref{artifactspresent} shows that theory
documentation is common.  We observed 17 of 24 packages had at least some theory
documentation. Templates for scientific requirements, like \citet{SmithEtAl2007,
SmithAndLai2005}, show theory is a significant part of the requirements
documentation. With the addition of some extra information, the theory documents
can be transformed into requirements specification.  The key extra information
includes explicitly stating user characteristics, explicitly stating how the
user interacts with the software in terms of input data requirements, and
listing likely and unlikely changes.  Explicit statements about likely changes
are invaluable in the design stage, since they provide developers guidance on
how general the software needs to be.  For those wishing to maximize the value
of a requirement specification, information could be added like prioritizing the
nonfunctional requirements (to show the relative importance between qualities
like portability, reliability, and performance) and traceability information (to
show the consequences of changes to the assumptions). Spending time to
incorporate additional information in the theory manuals should help address the
quality of existing documentation (\ppref{P_Documentation})  and improve the
software's usability (\ppref{P_Usability}).

@@

	\item[Repeatability] means producing the same output over successive runs of
	a program.  As was pointed out during our developer interviews, this can be
	difficult for LBM software because of dependence on probability
	distributions.  However, deterministic output is important for testing
	purposes and for debugging problematic code.  For LBM software repeatability
	will likely require running a serial version of the code with full control
	of the initialization of the pseudo-random number generators.

@@

DSM provides a transformative technology for documentation, design, and
verification \citep{JohansonAndHasselbring2018, Smith2018}. DSM allows
scientists to focus on their science, not software.  A generative approach
removes the maintenance nightmare of documentation duplicates and near
duplicates \citep{LucivEtAl2018}, since developers capture knowledge once and
transform it as needed.  Code generation has previously been applied to improve
research software.  For instance, ATLAS (Automatically Tuned Linear Algebra
Software) \citep{WhaleyEtAl2001} and Blitz++ \citep{Veldhuizen1998} produces
efficient and portable linear algebra software.  Spiral \citep{Pueschel2001}
uses software/hardware generation for digital signal processing.
\citet{Carette2008} shows how to generate a family of efficient, type-safe
Gaussian elimination algorithms.  FEniCS (Finite Element and Computational
Software) uses code generation when solving differential equations
\citep{LoggEtAl2012}. \citet{OberEtAl2018} and \citet{MatkerimEtAl2013} apply
DSM to High Performance Computing (HPC), using UML (Unified Modelling Language)
for their domain models. Unlike previous DSM work, the current recommendation
focuses on generating all software artifacts (requirements, design, etc.), not
just code.  \citet{SzymczakEtAl2016} presents initial work on this ``generate
all things'' approach and \citep{CaretteEtAl2021-Drasil} provides a prototype.

@@

\begin{enumerate}
	\item Employ linters to address \ppref{P_LackDevTime}, \ppref{P_LackSoftDevExp}, \ppref{P_TechnologyHurdles} and \ppref{P_TechDebt}. (Section~\ref{Sec_Linters})
	\item Conduct rigorous peer reviews to address \ppref{P_LackDevTime},
	\ppref{P_LackSoftDevExp}, \ppref{P_TechnologyHurdles} and
	\ppref{P_TechDebt}.	(Section~\ref{Sec_PeerReview})
	\item Write and submit papers on software to address \ppref{P_LackFunding}.
	(Section~\ref{Sec_MorePapers})
	\item Apply advice from the open-source community on how to grow the number
	of contributors to address \ppref{P_LackDevTime}.
	(Section~\ref{Sec_GrowContributors})
	\item Further address the oracle problem for testing, with techniques such as the method of manufactured solutions and metamorphic testing, to address
	\ppref{P_Correctness}. (Section~\ref{Sec_FurtherOracle})
	\item Augment theory manuals to include requirements information to address
	\ppref{P_LackDevTime}, \ppref{P_Usability} and \ppref{P_Documentation}.
	(Section~\ref{Sec_AugTheoryMans})
	\item Improve re-runnability, repeatability, reproducibility, and
	replicability to improve \ppref{P_Correctness}.
	(Section~\ref{Sec_Reproducibility})
	\item Use Domain Specific Modelling to generate all LBM code, build scripts,
	documentation, test cases from a stable knowledge base to address
	\ppref{P_LackDevTime}, \ppref{P_LackSoftDevExp},
	\ppref{P_TechnologyHurdles}, \ppref{P_Correctness}, \ppref{P_Usability},
	\ppref{P_TechDebt} and \ppref{P_Documentation}.
	(Section~\ref{Sec_GenAllThings})
\end{enumerate}

>>>>

\subsection{CI Usage Compared to Research Software in General} \label{Sec_CI}

@@

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are suggestions for consideration in the future. We
will not be repeating previously discussed ideas here, such as CI, documentation
of APIs, etc.  Our aim is to mention ideas that are at least somewhat beyond
conventional best practices. The ideas listed here have the potential to become
best practices in the medium to long-term. The ideas in the following
subsections are roughly in the order of increasing implementation effort.

@@

Except for \citet{ThielEtAl2020}, the research software guidelines that we
consulted do not mention linters.  However, we believe that linters have the
potential to improve code quality at a relatively low cost.  A linter is a tool
that statically analyzes code to find programming errors, stylistic
inconsistencies and suspicious constructs \citep{Wikipedia2022_Lint}. Linters
can be used to spot check code files, or even better as part of a continuous
integration system, as discussed in Section~\ref{Sec_CI}.

@@

Linters can address the pain points for LBM developers.  For instance, a linter
can decrease the amount of development time (\ppref{P_LackDevTime}) by
decreasing the number of mundane mistakes programmers have to catch.  Since the
linter can include rules that capture the wisdom of senior programmers, it can
help newer developers avoid common mistakes (\ppref{P_LackSoftDevExp}). Although
there are not many linters for checking parallel programming mistakes
(\ppref{P_TechnologyHurdles}), Parallel Lint is an option for OpenMP.
Consistently using a linter can guard against some technical debt
(\ppref{P_TechDebt}). Although linters are tools for code analysis, similar
ideas can be applied to ``lint'' documentation to ensure adherence to basic
rules.  The use of tools to check documentation is a partial explanation for the
relatively higher quality of statistical tools that are part of the
Comprehensive R Archive Network (CRAN) \citep{SmithEtAl2018_StatSoft}.

@@

Peer review is a strategy already adopted by some LBM developers to improve
confidence in their software (Section~\ref{SecReproducibility}), but the
benefits of peer review can be improved by making its application more rigorous.
\citet{Jones2008} shows that rigorous inspection finds 60-65\% of latent defects
on average, and often tops 85\% in defect removal efficiency.  By way of
comparison, most forms of testing average between 30 and 35\% for defect removal
efficiency \citep{EbertAndJones2009, Jones2008}.  A formal code inspection
involves asking the reviewer to either follow a review checklist (check
consistency of variable names, look for terminating loops, etc.), or perform
specific review tasks (like summarize the purpose of the code, create a data
dictionary for a given module, cross-reference the code to the technical
manual, etc.) The task based inspection approach has been effectively used for
research software, as described by \citet{KellyAndShepard2000}. Task based
inspection is an ideal fit with an issue tracking system, like GitHub.  The
review tasks can be issues, so that they can be easily assigned, monitored and
recorded.  Other potential issues for the tracker include assigning junior team
members to test installation instructions and getting-started tutorials.

Rigorous peer review addresses the same pain points as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_LackSoftDevExp},
\ppref{P_TechnologyHurdles}, and \ppref{P_TechDebt}. Peer review addresses these
points by efficiently searching for defects and problems.  Finding
misunderstandings in of how the code implements the required theory improves the
software's correctness (\ppref{P_Correctness}).

@@

To address the pain point of a lack of funding (\ppref{P_LackFunding}), LBM
developers should watch for new opportunities to publish their research software
source code. For instance, the Journal of Open Source Software (JOSS)
\citep{SmithEtAl2018-joss} reviews and publishes articles that credit the
scholarship contained in the software itself.  \citet{SmithEtAl2016-softcite}
presents a set of software citation principles that may encourage broad adoption
of a consistent policy for software citation across disciplines and venues.
\citet{ChueHongEtAl2019} provides a software citation checklist for developers
to make a release of their software citable.  \citet{KatzEtAl2021} provides
further guidance on software citation.  Another option is to use GitHub for
citation, since it is relatively easy to add metadata to repositories and to
generate citations for those repositories \citep{Smith2021-CitationsOnGitHub}.

@@

To address the pain point of ensuring correctness (\ppref{P_Correctness}),
Section~\ref{SecReproducibility} outlined techniques that LBM developers use to
circumvent the oracle problem for testing.  Additional techniques to build
confidence in correctness are as follows \citep{Smith2016}:

@@

As discussed in Section~\ref{Sec_CompareArtifacts}, developers rarely write
requirements specifications for research software. Although a full requirements
specification has advantages \citep{SmithEtAl2007, SmithAndLai2005}, given the
lack of development time (\ppref{P_LackDevTime}), we recommend the intermediate
step of augmenting the existing theory manuals with some requirements
information.  Table~\ref{artifactspresent} shows that theory documentation is
common.  We observed 17 of 24 packages had at least some theory documentation.
Templates for scientific requirements, like \citet{SmithEtAl2007,
SmithAndLai2005}, show theory is a significant part of the requirements
documentation. With the addition of some extra information, the theory documents
can be transformed into requirements specification.  The key extra information
includes explicitly stating user characteristics, explicitly stating how the
user interacts with the software in terms of input data requirements, and
listing likely and unlikely changes.  Explicit statements about likely changes
are invaluable in the design stage, since they provide developers guidance on
how general the software needs to be.  For those wishing to maximize the value
of a requirement specification, information could be added like prioritizing the
nonfunctional requirements (to show the relative importance between qualities
like portability, reliability, and performance) and traceability information (to
show the consequences of changes to the assumptions). Spending time to
incorporate additional information in the theory manuals should improve the
software's usability (\ppref{P_Usability}) and help address the quality of
existing documentation (\ppref{P_Documentation}).

@@

	\item[Repeatability] means producing the same output over successive runs of
	a program.  As pointed out during our developer interviews, this can be
	difficult for LBM software because of dependence on probability
	distributions.  However, deterministic output is important for testing
	purposes and for debugging problematic code.  For LBM software repeatability
	will likely require running a serial version of the code with full control
	of the initialization of the pseudo-random number generators.

@@

DSM provides a transformative technology for documentation, design, and
verification \citep{JohansonAndHasselbring2018, Smith2018}. DSM allows
scientists to focus on their science, not software.  A generative approach
removes the maintenance nightmare of documentation duplicates and near
duplicates \citep{LucivEtAl2018}, since developers capture knowledge once and
transform it as needed.  Code generation has previously been applied to improve
research software.  For instance, ATLAS (Automatically Tuned Linear Algebra
Software) \citep{WhaleyEtAl2001} and Blitz++ \citep{Veldhuizen1998} produces
efficient and portable linear algebra software.  Spiral \citep{Pueschel2001}
uses software/hardware generation for digital signal processing.
\citet{Carette2008} shows how to generate a family of efficient, type-safe
Gaussian elimination algorithms.  FEniCS (Finite Element and Computational
Software) uses code generation when solving differential equations
\citep{LoggEtAl2012}. \citet{OberEtAl2018} and \citet{MatkerimEtAl2013} apply
DSM to High Performance Computing (HPC), using UML (Unified Modelling Language)
for their domain models. Unlike previous DSM work, the current recommendation
focuses on generating all software artifacts (requirements, design, etc.), not
just code.  \citet{SzymczakEtAl2016} presents initial work on this ``generate
all things'' approach, \citet{SmithAndCarette2021-BRIC} presents a motivating
example, and \citep{CaretteEtAl2021-Drasil} provides a prototype.

@@

\begin{enumerate}
	\item Employ linters to address \ppref{P_LackDevTime}, \ppref{P_LackSoftDevExp}, \ppref{P_TechnologyHurdles} and \ppref{P_TechDebt}. (Section~\ref{Sec_Linters})
	\item Conduct rigorous peer reviews to address \ppref{P_LackDevTime},
	\ppref{P_LackSoftDevExp}, \ppref{P_TechnologyHurdles} and
	\ppref{P_TechDebt}.	(Section~\ref{Sec_PeerReview})
	\item Write and submit papers on software, to journals like JOSS, to address
	\ppref{P_LackFunding}. (Section~\ref{Sec_MorePapers})
	\item Apply advice from the open-source community on how to grow the number
	of contributors to address \ppref{P_LackDevTime}.
	(Section~\ref{Sec_GrowContributors})
	\item Further address the oracle problem for testing, with techniques such as the method of manufactured solutions and metamorphic testing, to address
	\ppref{P_Correctness}. (Section~\ref{Sec_FurtherOracle})
	\item Augment theory manuals to include requirements information to address
	\ppref{P_LackDevTime}, \ppref{P_Usability} and \ppref{P_Documentation}.
	(Section~\ref{Sec_AugTheoryMans})
	\item Improve re-runnability, repeatability, reproducibility, and
	replicability to improve \ppref{P_Correctness}.
	(Section~\ref{Sec_Reproducibility})
	\item Use Domain Specific Modelling to generate all LBM code, build scripts,
	documentation, test cases from a stable knowledge base to address
	\ppref{P_LackDevTime}, \ppref{P_LackSoftDevExp},
	\ppref{P_TechnologyHurdles}, \ppref{P_Correctness}, \ppref{P_Usability},
	\ppref{P_TechDebt} and \ppref{P_Documentation}.
	(Section~\ref{Sec_GenAllThings})
\end{enumerate}
